---
title: 'Reproducible Research: Peer Assessment 1'
output: html_document
---
## Loading and preprocessing the data

1. Load the the data into R.

```{r}
ActivityData = read.csv("activity.csv", colClasses = c("numeric", "character", "numeric"))

```

2. Check and transform the data into a format suitable for the analysis

- Checking the data load 
```{r}
head(ActivityData, 3)
tail(ActivityData, 3)
```

- Checking the data structure

```{r}
str(ActivityData)
```

- Summary of the data

```{r}
summary(ActivityData)
```

3. Transform the data

```{r}
ActivityData$date = as.Date(ActivityData$date, "%Y-%m-%d")
good =complete.cases(ActivityData)
ActivityData = ActivityData[good,]
```


## What is mean total number of steps taken per day?

1. Calculating the total number of steps taken per day

```{r}
library(ggplot2)

StepsDat = as.data.frame(sapply(split(ActivityData$steps, ActivityData$date), sum, na.rm=TRUE))
p = ggplot(StepsDat, aes(x=StepsDat[,1]))
p + geom_histogram() + labs(title= 'Total Steps by day') + xlab('Day') 

```

2. alculating and reporting the mean and median of the total number of steps taken per day

```{r}
mean(StepsDat[,1])
median(StepsDat[,1])
```

## What is the average daily activity pattern?

1. Making a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
library(ggplot2)
ts_Steps = sapply(split(ActivityData$steps, ActivityData$interval), mean, na.rm=TRUE)
ts_StepsDat = data.frame(days=as.numeric(names(ts_Steps)),ts_Steps) 
p = ggplot(ts_StepsDat, aes(x=days, y=ts_Steps))
p + geom_line(col='salmon')+labs(title='Average number of steps taken') +xlab('5-Min Interval') +ylab('Average across all days')
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
Max.Interval = which.max(ts_Steps)
names(Max.Interval)
```

## Imputing missing values

1. Calculating and reporting the total number of missing values in the dataset
```{r echo=FALSE}
ActivityData = read.csv("activity.csv", colClasses = c("numeric", "character", "numeric"))
ActivityData$date = as.Date(ActivityData$date, "%Y-%m-%d")
```

```{r}
sum(!complete.cases(ActivityData))
```

2. Devising a strategy for filling in all of the missing values in the dataset.
```{r}
AvgSteps = sapply(split(ActivityData$steps, ActivityData$interval), mean, na.rm=TRUE)
AvgSteps = data.frame(interval=as.numeric(names(AvgSteps)),AvgSteps) 

ActivityData = merge(ActivityData,AvgSteps, by = "interval")
NAs = is.na(ActivityData$steps)
ActivityData$steps[NAs] = ActivityData$AvgSteps[NAs]

```

3. Creating a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
ActivityDataNew = ActivityData[,c(1:3)]
head(ActivityDataNew)
tail(ActivityDataNew)
```
4. Making a histogram of the total number of steps taken each day and Calculating and reporting the mean and median total number of steps taken per day. 
```{r}
library(ggplot2)
StepsNewTot = sapply(split(ActivityDataNew$steps, ActivityDataNew$date), sum)
StepsNewDat = data.frame(StepsNewTot)

p <- ggplot(StepsNewDat, aes(x=StepsNewDat[,1]))
p + geom_histogram() + labs(title= 'Total Steps by day') + xlab('Day') 
```

```{r}
mean(StepsNewTot)
median(StepsNewTot)
```

## Are there differences in activity patterns between weekdays and weekends?

1. Creating a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day. 

```{r}
ActivityDataNew$date = as.Date(ActivityDataNew$date, format ='%Y-%m-%d')
ActivityDataNew$day =  ifelse(weekdays(ActivityDataNew$date) %in% c("sábado", "domingo"),'weekend','weekday')

StepsIntDay = aggregate(steps~ interval+day,ActivityDataNew,FUN="mean") 

```
2. Making a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.

```{r}
library(ggplot2)
p = ggplot(StepsIntDay, aes(x=interval,y=steps)) 
p + geom_line(col='salmon')+facet_grid(.~day)+ylab('Average Steps')+labs(title='Average steps pattern by weekday/weekend')
```




